Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2024-11-30 _tex/index.tex
Built site for gh-pages

For instance, Figure~\ref{fig-discriminal_process} illustrates the
discriminal process distributions along a quality trait continuum for
two written texts. The figure shows that these processes follow a Normal
distribution. Moreover, it depicts differences in the texts' positions
along the quality trait continuum, where text B is positioned two units
further along the continuum than text A, as indicated by their modal
discriminal processes (\(S_{B}=2\) and \(S_{A}=0\)). Finally, it
highlights differences in the texts' discriminal dispersions
(\(\sigma_{B}=1\) and \(\sigma_{A}=0.5\)), showing that text B exhibits
a greater variability in its perceived quality than text A, as reflected
by its wider distribution.

@@

Thus, the theory assumes that the observed dichotomous outcome arises
from the distribution of the difference between the underlying
discriminal processes of the stimuli, known as the \emph{discriminal
difference}. Since the individual discriminal processes follow a Normal
distribution, their difference also follows a Normal distribution. The
mode (or mean) of this distribution, representing the (average) relative
separation, is given by the difference between the modal discriminal
processes of the stimuli \(S_{BA} = S_{B} - S_{A}\). Meanwhile, the
dispersion of the distribution, reflecting the variability in the
relative separation, is calculated as
\(\sigma_{BA} = \sqrt{\sigma_{B}^{2} + \sigma_{A}^{2} - \rho\sigma_{B}\sigma_{A}}\).
Here, \(\sigma_{B}\) and \(\sigma_{A}\) are the previously defined
discriminal dispersions, while \(\rho\) measures the correlation between
their discriminal processes. This correlation quantifies how much the
judge's perception of the trait in one stimulus influences his
perception of the same trait in the other.

@@

For instance, when two texts clearly differ in quality, as depicted in
Figure~\ref{fig-discriminal_process}, higher correlations increase the
likelihood that the discriminal difference more distinctly favors text B
over text A, as indicated by the shaded gray area under the curve in
Figure~\ref{fig-correlation}. Conversely, it is easy to infer that when
the two texts have similar or identical quality levels, higher
correlations reduce the likelihood that the discriminal difference
distinctly favors one text over the other, as the distribution narrows
around zero (not depicted in the figure).

>>>>

For instance, Figure~\ref{fig-discriminal_process} illustrates the
discriminal process distributions along a quality trait continuum for
two written texts. The figure shows that these processes follow a Normal
distribution. Moreover, it depicts differences in the texts' positions
along the quality trait continuum, where text B is positioned further
along the continuum than text A, as indicated by their modal discriminal
processes (\(S_{B}\) and \(S_{A}\)). Finally, it highlights differences
in the texts' discriminal dispersions (\(\sigma_{B}\) and
\(\sigma_{A}\)), showing that text B exhibits a greater variability in
its perceived quality than text A, as reflected by its wider
distribution.

@@

Indeed, the theory assumes that the observed dichotomous outcome arises
from the distribution of the difference between the underlying
discriminal processes of the stimuli, known as the \emph{discriminal
difference}. Since the individual discriminal processes follow a Normal
distribution, their difference also follows a Normal distribution. The
mode (or mean) of this distribution, representing the (average) relative
separation, is given by the difference between the modal discriminal
processes of the stimuli \(S_{BA} = S_{B} - S_{A}\). Meanwhile, the
dispersion of the distribution, reflecting the variability in the
relative separation, is calculated as
\(\sigma_{BA} = \sqrt{\sigma_{B}^{2} + \sigma_{A}^{2} - \rho\sigma_{B}\sigma_{A}}\).
Here, \(\sigma_{B}\) and \(\sigma_{A}\) are the previously defined
discriminal dispersions, while \(\rho\) measures the correlation between
their discriminal processes. This correlation quantifies how much the
judge's perception of the trait in one stimulus influences his
perception of the same trait in the other.

@@

For instance, when two texts clearly differ in quality, as depicted in
Figure~\ref{fig-discriminal_process}, higher correlations increase the
likelihood that the discriminal difference more distinctly favors text B
over text A, as indicated by the shaded gray area under the curve in
Figure~\ref{fig-correlation}. Conversely, when the two texts have
similar or identical quality levels, higher correlations reduce the
likelihood that the discriminal difference distinctly favors one text
over the other, as the distribution narrows around zero (not depicted in
the figure).
